---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Jitterbox

<!-- badges: start -->
<!-- badges: end -->

This is a package created by Brynley Hanson-Wright for the course STAT 545B taught by 
Vincenzo Coia in 2021. The goal of this package is to allow you to make quick plots that compare the distribution of a numeric variable within groups of a categorical variable. It displays boxplots, stripplots, and means within each group to visualize both raw data and summary statistics.

## Installation

Jitterbox is not on CRAN. You can download the package from this repository with:
  
```{r}
#install.packages("devtools")
#devtools::install_github("brynhw/Jitterbox") 
```
  
## Examples  

One simple example is looking at Edgar Anderson's Iris dataset and plotting the distribution of sepal length for each species.  

```{r}
library(Jitterbox)
box.bygroup(iris, Sepal.Length, Species, numeric.lab = "Sepal length", group.lab = "Iris species")
```

Another example with the ToothGrowth dataset shows how the function can use non-factor inputs for the grouping variable. In this case dose is numeric, but given there are three levels of dose it makes sense to treat these as categorical levels. The function will do this and return a warning in case a non-factor variable was mistakenly entered.  

```{r, warning = TRUE}
box.bygroup(ToothGrowth, len, dose)  
```

Lastly, here is an example using life expectancy data by continent from the Gapminder dataset. Because there is quite a bit more data in this plot, the alpha level of the points can be adjusted with the point.alpha argument.

```{r}
box.bygroup(gapminder::gapminder, lifeExp, continent, point.alpha = 0.3)  
```
